---
title: "BRP"
author: "Neeba E A"
date: "July 06, 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BRP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BRP)
```


This package is primarily built for educational purposes. The package enables one to explore fatal traffic accidents from 2013-2015 using data from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (BRP).  

## The Data

The data in this package come from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (BRP) data.

```{r sample2013_data}
BRP_2013_fn <- make_filename(2013)
BRP_2013_fn
BRP_2013 <- BRP_read(BRP_2013_fn) 
dim(BRP_2013)
BRP_2013
```

For detailed information about the data, see the [NHTSA BRP Manuals & Documentation page](https://crashstats.nhtsa.dot.gov/#/DocumentTypeList/4). 

## Loading BRP Data

To load all of the data for a given year, use the `make_filename()` and `BRP_read()` functions, as shown in the previous section.

### About the Filename

Use the `make_filename` command to find out where data is stored on your machine and/or create a filename to save/load new data.

BEWARE: re-installing the package may cause your data to be overridden

```{r make_filename}
BRP_2013_fn <- make_filename(2013)
BRP_2013_fn
```

### Single Year

If you wish to just look at fatality data for a a single year, use the `BRP_read_years()` function with a single year as input. The only data columns selected are `MONTH` and `year`.  This returns a list of length one, and the first element in the list is the `tbl_df` (the `tidyverse` data frame) listing the month and year for each fatal accident.

```{r single_year}
BRP_2014 <- BRP_read_years(years = 2014)
BRP_2014[[1]]
```

### Multiple Years

If you wish to look at fatalities for multiple years, enter a vector of years as the argument for the `BRP_read_years()` function (examples: `BRP_read_years(years = c(2013, 2015))` or `BRP_read_years(2013:2015)`.  Again, this returns a list of `tbl_df`s,  with each element of the list showing the month and year for each fatality. 

```{r multiple_years}
BRP_3yrs <- BRP_read_years(years = 2013:2015)
BRP_3yrs
```

## Summarizing BRP Data

The `BRP_summarize_years()` function take the same argument as the `BRP_read_years()`, and produces a summary of the simple counts of fatalities by month and year: 

```{r summarize_data}
BRP_summary <- BRP_summarize_years(2013:2015)
BRP_summary
```


## Mapping Fatal Crashes

Finally, the `BRP_map_state` function takes a state ID number and a year, and maps that state's fatalities with a dot at the fatality location.  Note that in order to use this function, you will likely need to load the `mapdata` package.  

For a list of the state ID numbers, see page 26 of the [BRP Analytical User's Guide (2015)](https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812315#27). 

```{r mapping_crashes}
library(mapdata)
BRP_map_state(53, 2014)
BRP_map_state(36, 2014)
```
